---
title: "Homework 2"
author: "Sarah McLarnan"
date: 2019-10-03
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
```

### Problem 1 
```{r problem_1_water_wheel, message = FALSE}
water_wheel = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, skip = 1) %>%
  janitor::clean_names() %>%
  select(-x15, -x16, -x17) %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(round(sports_balls)))
```
```{r problem_1_precip_2017}
precip_2017 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017)
```

```{r problem_1_precip_2018}
precip_2018 =
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 6, skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018)
```
```{r problem_1_precip_tidy}
precip_tidy =
  bind_rows(precip_2017, precip_2018) %>%
  mutate(month = month.name[month])
```
There are `r nrow(precip_tidy)` observations in the precipitation data set and `r nrow(water_wheel)` observations in the water wheel data set. The water wheel dataset has information on the dumpster materials were collected from, the date of collection, the amount of materail collected in various categories (i.e. `plastic_bottles`, `chip_bages`, etc.) and how many homes were powered with the materials collected. The precipitation data includes columns for the year, month and total precipitation. The total precipitation in 2018 was `r sum(pull(filter(precip_tidy, year == "2018"), total))` inches. In 2017 the median number of sports balls in dumpsters was `r median(pull(filter(water_wheel, year == 2017), sports_balls))`. 

### Problem 2 

```{r problem_2_pols_month}
pols_month = 
  read_csv("./data/pols-month.csv") %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(year = as.integer(year),
  month = as.integer(month),
  day = as.integer(day),
  month = month.name[month], 
  president = prez_dem,
  president = recode(president, "0" =  "gop", "1" = "dem")) %>%
  select(-prez_gop, -prez_dem, -day)
```

```{r problem_2_snp}
snp = 
  read_csv("./data/snp.csv") %>%
  separate(date, into = c("month", "day", "year")) %>%
  select (year, month, close, -day) %>%
  mutate(year = as.integer(year),
  month = as.integer(month),
  month = month.name[month])
```
```{r problem_2_unemployment}
unemployment =
  read_csv("./data/unemployment.csv") %>%
  janitor::clean_names()
  unemployment = pivot_longer(unemployment, jan:dec, names_to = "month", values_to = "prec_unemployment") %>%
  mutate(month = recode(month, 
                        "jan" = "January", 
                        "feb" = "February", 
                        "mar" = "March", 
                        "apr" = "April", 
                        "jun" = "June", 
                        "jul" = "July", 
                        "aug" = "August", 
                        "sep" = "September", 
                        "oct" = "October", 
                        "nov" = "November", 
                        "dec" = "December"))
```
```{r problem_2_data_merge}
pols_snp = left_join(pols_month, snp, by = c("year", "month"))
pols_snp_une = left_join(pols_snp, unemployment, by = c("year", "month"))
```
The `pols_month` data set has `r nrow(pols_month)` observations of `r ncol(pols_month)` variables incluiding the year, month, number of gop and democratic governors, senators, and representatives and affilation of the president. The `snp` data set has `r nrow(snp)` observations of `r ncol(snp)` variables including the month and year and the closing values of the S&P stock index. The `unemployment` data set has `r nrow(unemployment)` observations of `r ncol(unemployment)` variables which including the month, year and percent unemployment. The merged data set has `r nrow(pols_snp_une)` and `r ncol(pols_snp_une)` variables. 

### Problem 3
```{r problem_3_baby_names}
baby_names =
  read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  distinct() %>%
  mutate(gender = tolower(gender), 
        ethnicity = tolower(ethnicity),
        ethnicity = recode(ethnicity, 
                       "asian and pacific islander" = "asian and paci", 
                       "black non hispanic" = "black non hisp", 
                       "white non hispanic" = "white non hisp"))
```
```{r problem_3_olivia}
olivia = filter(baby_names, childs_first_name == "Olivia") %>%
  select(-gender, -childs_first_name, -count) %>%
  pivot_wider(names_from = "year_of_birth", values_from = "rank")

knitr::kable(olivia)
```
```{r problem_3_popular_male}
popular_male = filter(baby_names, gender == "male", rank == "1") %>%
  select(-gender, -rank, -count) %>%
  pivot_wider(names_from = "year_of_birth", values_from = "childs_first_name")

knitr::kable(popular_male)
```
```{r problem_3_white_males_graph}
white_males_2016 = filter(baby_names, gender == "male" & year_of_birth == "2016" & ethnicity == "white non hisp")

ggplot(white_males_2016, aes(x = rank, y = count)) +
  geom_point()
```





